Supplementary Information -- Morphological convergence as on-line lexical analogy

### 2020-02-19 supplementary_information/si.Rnw
mgl sum typo

## get the raw output of the baseline feature-distance mgl
mgl_baseline_rules_raw = readr::read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
  dplyr::mutate(
    form1 = stringr::str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
    form2 = stringr::str_replace_all(form2, '»', '')
  )

@@

>>>>

## get the raw output of the baseline feature-distance mgl
mgl_baseline_rules_raw = readr::read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
  dplyr::mutate(
    form1 = stringr::str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
    form2 = stringr::str_replace_all(form2, '»', ''),
    form2 = stringr::str_replace_all(form2, 'o', '5') # transcription thing
  )

@@

# verb_handles

### 2020-02-19 supplementary_information/si.tex
mgl sum typo

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 18 16:57:29 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllll}
  \hline
 & verb class & irregular set & regular set & miscellaneous regular verbs \\ 
  \hline
1 & BURNT & 6 & 42 & 1218 \\ 
  2 & DROVE & 14 & 83 & 1218 \\ 
  3 & KEPT & 12 & 31 & 1218 \\ 
  4 & SANG & 8 & 13 & 1218 \\ 
  5 & sum & 40 & 169 & 1218 \\ 
   \hline
\end{tabular}
\caption{Number of forms in each verb class, GCM training data} 
\label{gcmsetsize}
\end{table}
\subsection{Estimation}

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 18 16:57:29 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllrr}
  \hline
 & category & rule.type & related forms & exceptions \\ 
  \hline
1 & burnt & irregular &  24 &  41 \\ 
  2 & kept & irregular &  10 &   9 \\ 
  3 & sang & irregular &  11 &  77 \\ 
  4 & burnt & regular &  38 &  21 \\ 
  5 & drove & regular &  29 &  27 \\ 
  6 & kept & regular &  67 &  41 \\ 
  7 & sang & regular &  47 &  38 \\ 
   \hline
\end{tabular}
\caption{Number of forms in each verb class, MGL training data} 
\label{mglsetsize}
\end{table}
\subsection{Estimation}

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 18 16:57:29 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllrr}
  \hline
 & category & rule.type & related forms & exceptions \\ 
  \hline
1 & burnt & irregular &  24 &  41 \\ 
  2 & kept & irregular &  10 &   9 \\ 
  3 & sang & irregular &  11 &  77 \\ 
  4 & burnt & regular &  38 &  21 \\ 
  5 & drove & regular &  29 &  27 \\ 
  6 & kept & regular &  67 &  41 \\ 
  7 & sang & regular &  47 &  38 \\ 
   \hline
\end{tabular}
\caption{The number of unique rules posited by the MGL in each verb class} 
\label{mglruleses}
\end{table}
Table \ref{mglruleseskept} shows the list of (irregular and regular) rules for one of the classes, the KEPT class. We choose this class as it provides a good illustration of the estimation process. An MGL rule describes an $A \rightarrow{} B$ {\em alternation} in an $XAY$ {\em environment}. Since the KEPT rules are all suffixing, the following environment ($Y$) is empty. For the preceding environment ($X$), we give a list of segments that are matched by the structural description if it applies to four segments or fewer. Otherwise, we give a phonological description. So, for instance, the first rule turns \textipa{[il]} to \textipa{[Elt]} following any obstruent. Note that in some cases, the MGL generates multiple rules for the same alternation (such as \textipa{ip} $\rightarrow{}$ \textipa{Ept}). As discussed in Section \ref{mgloutput}, in such instances, we choose the candidate rule with the highest adjusted confidence in order to calculate the MGL regularity score. (Fricatives and nasal consonants constitute a natural class in the segmental feature system used by \citealt{albright2003rules}.)

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 18 16:57:29 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -0.108 & 0.161 & -0.672 & 0.501 \\ 
  verb\_baseline\_average & 8.259 & 0.686 & 12.040 & 0.000 \\ 
  bot\_lexical\_typicalityrandom & 0.043 & 0.169 & 0.253 & 0.800 \\ 
  bot\_lexical\_typicalityreversed & 0.035 & 0.173 & 0.204 & 0.839 \\ 
  participant\_prereg\_average & 7.125 & 0.563 & 12.667 & 0.000 \\ 
  bot\_regularisation\_rate-40\% & -0.456 & 0.185 & -2.468 & 0.014 \\ 
  bot\_regularisation\_rate+40\% & 0.549 & 0.166 & 3.307 & 0.001 \\ 
  verb\_baseline\_average:bot\_lexical\_typicalityrandom & -2.402 & 0.927 & -2.591 & 0.010 \\ 
  verb\_baseline\_average:bot\_lexical\_typicalityreversed & -2.127 & 0.955 & -2.226 & 0.026 \\ 
   \hline
\end{tabular}
\caption{Experimental Data, Post-test Regression Model Summary. Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)} 
\label{regressiontable1}
\end{table}
Our regression model selection in this article proceeded as follows. Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}. Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function. However, our model tables (e.g., Tables 5 and 6 in the main text) include the significance levels generated by lme4 summaries.

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 18 16:57:30 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -0.525 & 0.124 & -4.219 & 0.000 \\ 
  rescaled\_CELEX\_mgl\_feature\_distance & 1.347 & 0.093 & 14.482 & 0.000 \\ 
  rescaled\_CELEX\_gcm\_feature\_distance\_residual & 1.423 & 0.128 & 11.100 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model 3: MGL scores and GCM residuals, baseline data} 
\label{resid1}
\end{table}
% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 18 16:57:31 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -1.225 & 0.160 & -7.682 & 0.000 \\ 
  rescaled\_CELEX\_gcm\_feature\_distance & 2.391 & 0.149 & 16.063 & 0.000 \\ 
  rescaled\_CELEX\_mgl\_feature\_distance\_residual & 1.008 & 0.090 & 11.154 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model 4: GCM scores and MGL residuals, baseline data} 
\label{resid2}
\end{table}
Model 1 predicts the GCM scores from the MGL scores (M1: GCM regular score  $\sim$ MGL regular score). Model 2 predicts the MGL scores from the GCM scores (M2: MGL regular score  $\sim$ GCM regular score). In two additional models, we then use the residuals from each model (M1, M2) as an estimate of the variation present in one score that cannot be attributed to the other.  First, we use the residuals of Model 1 in another, mixed-effects logistic regression model along with the MGL scores to predict regular responses in the baseline task (M3: regular response  $\sim$ MGL score + GCM residual + (1 + MGL score + GCM residual$|$participant) + (1 $|$ verb)). We then residualize for the MGL. We use the residuals of Model 2 along with the GCM scores in a similar logistic regression model (M4: regular response  $\sim$ GCM score + MGL residual + (1 + GCM score + MGL residual $|$participant)).

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 18 16:57:31 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -2.951 & 0.295 & -10.007 & 0.000 \\ 
  rescaled\_CELEX\_gcm\_feature\_distance & 4.369 & 0.425 & 10.284 & 0.000 \\ 
  gcm\_score\_diff & 2.612 & 0.470 & 5.563 & 0.000 \\ 
  participant\_prereg\_average & 5.289 & 0.602 & 8.788 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model 5: Celex-GCM and individual score difference, post-test data} 
\label{indiv1}
\end{table}
% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 18 16:57:31 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -1.318 & 0.152 & -8.663 & 0.000 \\ 
  rescaled\_CELEX\_mgl\_feature\_distance & 1.944 & 0.176 & 11.045 & 0.000 \\ 
  mgl\_score\_diff & 0.304 & 0.139 & 2.181 & 0.029 \\ 
  participant\_prereg\_average & 7.203 & 0.591 & 12.190 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model 6: Celex-MGL and individual score difference, post-test data} 
\label{indiv2}
\end{table}
We do this for both the GCM  (Model 5: regular response  $\sim$ \textsc{Celex} GCM score + extra GCM information + participant pre-test mean + (1 + \textsc{Celex} GCM + extra GCM information$|$participant) + (1$|$verb) and the MGL (Model 6: regular response  $\sim$ \textsc{Celex} MGL score + extra MGL information + participant pre-test mean + (1 + \textsc{Celex} MGL + extra MGL information$|$participant)+ (1$|$verb)).

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 18 16:57:31 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -2.942 & 0.270 & -10.896 & 0.000 \\ 
  r\_baseline\_mgl\_features & 1.600 & 0.171 & 9.361 & 0.000 \\ 
  r\_individual\_gcm\_features & 2.786 & 0.378 & 7.369 & 0.000 \\ 
  participant\_prereg\_average & 5.437 & 0.585 & 9.298 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Celex-MGL and Individual-GCM predict variation in the post-test} 
\label{comp3}
\end{table}
Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ 201.86) or individual-GCM ($\chi^2 = $ 81.82) as predictors. It is not significantly better if we include individual-MGL ($\chi^2 = $ 0.09) or {\sc Celex}-GCM ($\chi^2 = $ 2.03).

>>>>

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Feb 19 11:30:14 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllll}
  \hline
 & verb class & irregular set & regular set & miscellaneous regular verbs \\ 
  \hline
1 & BURNT & 6 & 42 & 1218 \\ 
  2 & DROVE & 14 & 83 & 1218 \\ 
  3 & KEPT & 12 & 31 & 1218 \\ 
  4 & SANG & 8 & 13 & 1218 \\ 
  5 & sum & 40 & 169 & 1218 \\ 
   \hline
\end{tabular}
\caption{Number of forms in each verb class, GCM training data} 
\label{gcmsetsize}
\end{table}
\subsection{Estimation}

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Feb 19 11:30:14 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllrr}
  \hline
 & category & rule.type & related forms & exceptions \\ 
  \hline
1 & burnt & irregular &  24 &  41 \\ 
  2 & drove & irregular &  21 & 114 \\ 
  3 & kept & irregular &  10 &   9 \\ 
  4 & sang & irregular &  11 &  77 \\ 
  5 & burnt & regular &  38 &  21 \\ 
  6 & drove & regular &  29 &  27 \\ 
  7 & kept & regular &  67 &  41 \\ 
  8 & sang & regular &  47 &  38 \\ 
   \hline
\end{tabular}
\caption{Number of forms in each verb class, MGL training data} 
\label{mglsetsize}
\end{table}
\subsection{Estimation}

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Feb 19 11:30:14 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllrr}
  \hline
 & category & rule.type & related forms & exceptions \\ 
  \hline
1 & burnt & irregular &  24 &  41 \\ 
  2 & drove & irregular &  21 & 114 \\ 
  3 & kept & irregular &  10 &   9 \\ 
  4 & sang & irregular &  11 &  77 \\ 
  5 & burnt & regular &  38 &  21 \\ 
  6 & drove & regular &  29 &  27 \\ 
  7 & kept & regular &  67 &  41 \\ 
  8 & sang & regular &  47 &  38 \\ 
   \hline
\end{tabular}
\caption{The number of unique rules posited by the MGL in each verb class} 
\label{mglruleses}
\end{table}
Table \ref{mglruleseskept} shows the list of (irregular and regular) rules for one of the classes, the KEPT class. We choose this class as it provides a good illustration of the estimation process. An MGL rule describes an $A \rightarrow{} B$ {\em alternation} in an $XAY$ {\em environment}. Since the KEPT rules are all suffixing, the following environment ($Y$) is empty. For the preceding environment ($X$), we give a list of segments that are matched by the structural description if it applies to four segments or fewer. Otherwise, we give a phonological description. So, for instance, the first rule turns \textipa{[il]} to \textipa{[Elt]} following any obstruent. Note that in some cases, the MGL generates multiple rules for the same alternation (such as \textipa{ip} $\rightarrow{}$ \textipa{Ept}). As discussed in Section \ref{mgloutput}, in such instances, we choose the candidate rule with the highest adjusted confidence in order to calculate the MGL regularity score. (Fricatives and nasal consonants constitute a natural class in the segmental feature system used by \citealt{albright2003rules}.)

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Feb 19 11:30:14 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -0.108 & 0.161 & -0.672 & 0.501 \\ 
  verb\_baseline\_average & 8.259 & 0.686 & 12.040 & 0.000 \\ 
  bot\_lexical\_typicalityrandom & 0.043 & 0.169 & 0.253 & 0.800 \\ 
  bot\_lexical\_typicalityreversed & 0.035 & 0.173 & 0.204 & 0.839 \\ 
  participant\_prereg\_average & 7.125 & 0.563 & 12.667 & 0.000 \\ 
  bot\_regularisation\_rate-40\% & -0.456 & 0.185 & -2.468 & 0.014 \\ 
  bot\_regularisation\_rate+40\% & 0.549 & 0.166 & 3.307 & 0.001 \\ 
  verb\_baseline\_average:bot\_lexical\_typicalityrandom & -2.402 & 0.927 & -2.591 & 0.010 \\ 
  verb\_baseline\_average:bot\_lexical\_typicalityreversed & -2.127 & 0.955 & -2.226 & 0.026 \\ 
   \hline
\end{tabular}
\caption{Experimental Data, Post-test Regression Model Summary. Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)} 
\label{regressiontable1}
\end{table}
Our regression model selection in this article proceeded as follows. Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}. Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function. However, our model tables (e.g., Tables 5 and 6 in the main text) include the significance levels generated by lme4 summaries.

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Feb 19 11:30:16 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -0.525 & 0.124 & -4.219 & 0.000 \\ 
  rescaled\_CELEX\_mgl\_feature\_distance & 1.347 & 0.093 & 14.482 & 0.000 \\ 
  rescaled\_CELEX\_gcm\_feature\_distance\_residual & 1.423 & 0.128 & 11.100 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model 3: MGL scores and GCM residuals, baseline data} 
\label{resid1}
\end{table}
% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Feb 19 11:30:16 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -1.225 & 0.160 & -7.682 & 0.000 \\ 
  rescaled\_CELEX\_gcm\_feature\_distance & 2.391 & 0.149 & 16.063 & 0.000 \\ 
  rescaled\_CELEX\_mgl\_feature\_distance\_residual & 1.008 & 0.090 & 11.154 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model 4: GCM scores and MGL residuals, baseline data} 
\label{resid2}
\end{table}
Model 1 predicts the GCM scores from the MGL scores (M1: GCM regular score  $\sim$ MGL regular score). Model 2 predicts the MGL scores from the GCM scores (M2: MGL regular score  $\sim$ GCM regular score). In two additional models, we then use the residuals from each model (M1, M2) as an estimate of the variation present in one score that cannot be attributed to the other.  First, we use the residuals of Model 1 in another, mixed-effects logistic regression model along with the MGL scores to predict regular responses in the baseline task (M3: regular response  $\sim$ MGL score + GCM residual + (1 + MGL score + GCM residual$|$participant) + (1 $|$ verb)). We then residualize for the MGL. We use the residuals of Model 2 along with the GCM scores in a similar logistic regression model (M4: regular response  $\sim$ GCM score + MGL residual + (1 + GCM score + MGL residual $|$participant)).

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Feb 19 11:30:16 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -2.951 & 0.295 & -10.007 & 0.000 \\ 
  rescaled\_CELEX\_gcm\_feature\_distance & 4.369 & 0.425 & 10.284 & 0.000 \\ 
  gcm\_score\_diff & 2.612 & 0.470 & 5.563 & 0.000 \\ 
  participant\_prereg\_average & 5.289 & 0.602 & 8.788 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model 5: Celex-GCM and individual score difference, post-test data} 
\label{indiv1}
\end{table}
% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Feb 19 11:30:17 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -1.318 & 0.152 & -8.663 & 0.000 \\ 
  rescaled\_CELEX\_mgl\_feature\_distance & 1.944 & 0.176 & 11.045 & 0.000 \\ 
  mgl\_score\_diff & 0.304 & 0.139 & 2.181 & 0.029 \\ 
  participant\_prereg\_average & 7.203 & 0.591 & 12.190 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model 6: Celex-MGL and individual score difference, post-test data} 
\label{indiv2}
\end{table}
We do this for both the GCM  (Model 5: regular response  $\sim$ \textsc{Celex} GCM score + extra GCM information + participant pre-test mean + (1 + \textsc{Celex} GCM + extra GCM information$|$participant) + (1$|$verb) and the MGL (Model 6: regular response  $\sim$ \textsc{Celex} MGL score + extra MGL information + participant pre-test mean + (1 + \textsc{Celex} MGL + extra MGL information$|$participant)+ (1$|$verb)).

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Feb 19 11:30:17 2020
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -2.942 & 0.270 & -10.896 & 0.000 \\ 
  r\_baseline\_mgl\_features & 1.600 & 0.171 & 9.361 & 0.000 \\ 
  r\_individual\_gcm\_features & 2.786 & 0.378 & 7.369 & 0.000 \\ 
  participant\_prereg\_average & 5.437 & 0.585 & 9.298 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Celex-MGL and Individual-GCM predict variation in the post-test} 
\label{comp3}
\end{table}
Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ 201.86) or individual-GCM ($\chi^2 = $ 81.82) as predictors. It is not significantly better if we include individual-MGL ($\chi^2 = $ 0.09) or {\sc Celex}-GCM ($\chi^2 = $ 2.03).
